Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-05-12 Manuscript/proto-manuscript.tex
Restructured text in proto-MS, now it follows only the main message with a few excerpts from Samuli's writing.

\textbf{Specific introduction: MDEC and current FFs}

 Current force fields -- pros and cons, at a good shape in many aspects, agree on various properties

 lipid force fields fail in description of membrane-cation interaction -- could be answered by ECC?

@@

How I apply ECC on Lipid14 POPC and why such choice (i.e. $f_\sigma$, $f_q$ and definition of the scaled region).

why Lipid 14 -- good ratio of $\alpha / \beta$ response.

\textbf{Observations:}

Accurate electrometer response to \ce{Na^+}, \ce{Ca^{2+}} and \ce{DHMDMAB} cationic surfactant.

Preserves other relevant properties: APL, Order Parameters (OPs), diffusion? \todoi{diffusion can be even improved}, elasticity? \todoi{check elastic properties -- do they change, improve, worsen??} -- however, it is still garbage in-garbage out

Gives correct scattering form factor -- good groud for interpretting experimental data.

Provides correct stoichiometry for \ce{Na^+}, \ce{Ca^{2+}} and their interaction energies with the lipid membrane.

@@

questionable charge distribution from RESP -- a leeway for further tweaks of the FF

acheivable accuracy of the MD engines themselves (mainly Hector's worry).

application of the correction to other lipid models: The rule looks general, however, it depends on how accurate the original model was.

@@

Say what we did...

@@

%\label{}
\section{Introduction}

Cations were shown to generally overbind in PC lipid bilayers in
NMRlipids II project. Here we propose that the cation overbinding can
be corrected by implicitly inducing electronic polarizability in
lipid headgroups by scaling the partial charges.

%\section{Methods}

%\subsection{Molecular dynamics simulations}

\section{Effect of scaling of headgroup and glycerol backbone partial charges on lipid bilayer properties and cation binding}

Headgroup and glycerol backbone partial charges of Slipids and Lipid14 models were scaled
with various scaling factors and cation binding was monitored by using the electrometer
concept. Headgroup order parameter changes as a function of ion concentration are
shown in Fig. \ref{OrderParameterIONSchangesSCALED}. The order parameter decrease
seemst to be proportional to the headgroup charges: the smaller the charges
(larger scaling factor), the less change is observed in order parameters.
According to the electrometer concept, this means that the headgroup scaling
decreases cation binding affinity. However, from this data only we cannot
exclude the possibility that order parameters are changing less because
headgroup becomes less sensitive for cation binding.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Fig/OrderParameterIONSchangesSCALED.eps}
  \caption{\label{OrderParameterIONSchangesSCALED}
    Headgroup order parameter changes as a function of cations from models
    with modified headgroup partial charges.}
\end{figure}

The best models from Fig.  \ref{OrderParameterIONSchangesSCALED} are shown in
Fig. \ref{OrderParameterCHANGESnewMODELS}. Based on these results we choose
two models for more careful studies: Slipids and Lipid 14 models with headroup
and glycerol backbone partial charges scaled with 0.8 and 0.85, respectively.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Fig/OrderParameterCHANGESnewMODELS.eps}
  \caption{\label{OrderParameterCHANGESnewMODELS}
    Headgroup order parameter changes as a function of cations from models
    with modified headgroup partial charges.}
\end{figure}

While scaling seems to improve ion binding, it changes the lipid bilayer properties
without ions. The glycerol backbone and headgroup order parameters are plotted in
Fig \ref{HGopsNEWmodels.eps} for reference models and EEC models. The difference,
i.e. the effect of scaling on these parameters, is shown in Fig. \ref{HGopsNEWmodelsCHANGE}.
The area per lipid from different models are shown in Table \ref{apls}
The conclusion is that the charge scaling has only a marginal effect on
headgroup order parameters, while area per lipid is significantly decreased.
Area per lipid decrease is more pronounced in Lipid14 than in Slipids.

@@

\begin{table}
  \caption{Area per lipid from different models for POPC without ions\label{apls} }
  \begin{tabular}{c c}
    model          & A (Å$^2$)    \\
    Lipid14 (literature)  & 65.6$\pm$ 0.5   \\
    Lipid14eec0.85 & 55.5      \\
    Slipids (literature T=303K)       & 64.6 $\pm$ 0.4   \\
    SlipidsEEC0.8  & 57.8  \\
  \end{tabular}
\end{table}

The Calcium densities from new models and standard models are compared in Fig \ref{CAdensities}.
It seems to me that Ca binding in Lipid14 model is not reduced by scaling the
partial charges even thought the order parameter response is better.
On the other hand, in Slipids the binding seems to be reduced.
Possible interpretation is that Lipid14 result gets better because headgroup
gets less sensitive for bound charge, while in Slipid the binding is reduced.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Fig/CAdensities.eps}
  \caption{\label{CAdensities}
    Calcium densities from different simulations with new and standard models.
  }
\end{figure}

\section{Conclusions}

>>>>

\textbf{Specific introduction: current FFs and MDEC/ECC}

 Current force fields -- pros and cons, at a good shape in many aspects, agree on various properties.

 lipid force fields fail in description of membrane-cation interaction -- could be answered by ECC? 
Cations were shown to generally overbind in PC lipid bilayers in
NMRlipids II project. Here we propose that the cation overbinding can
be corrected by implicitly inducing electronic polarizability in
lipid headgroups by scaling the partial charges.

@@

How do I apply ECC on Lipid14 POPC and why such choice (i.e. $f_\sigma$, $f_q$ and definition of the scaled region).

why did I choose Lipid 14 -- good ratio of $\alpha / \beta$ response.

\textbf{Observations, how does the system behave, what are its macroscopic properties:}

Accurate electrometer response to \ce{Na^+}, \ce{Ca^{2+}} and \ce{DHMDMAB} cationic surfactant. 
The response of the headgroup order parameters is improved both through diminished affinity towards the cations and through lower headgroup response to the bound charge (from simulations with \ce{DHMDMAB} surfactant). 
The best models from Fig.  \ref{OrderParameterIONSchangesSCALED} are shown in
Fig. \ref{OrderParameterCHANGESnewMODELS}. Based on these results we choose

Preserves other relevant properties: APL, Order Parameters (OPs), diffusion? \todoi{diffusion can be even improved}, elasticity? \todoi{check elastic properties -- do they change, improve, worsen??} -- however, it is still garbage in-garbage out
The glycerol backbone and headgroup order parameters are plotted in
Fig \ref{HGopsNEWmodels.eps} for reference models and EEC models. The difference,
i.e. the effect of scaling on these parameters, is shown in Fig. \ref{HGopsNEWmodelsCHANGE}.

Gives correct scattering form factor -- good ground for interpretting experimental data.

Provides/predicts correct stoichiometry for \ce{Na^+}, \ce{Ca^{2+}} and their interaction energies with the lipid membrane.

@@

questionable charge distribution from RESP -- a leeway for further tweaks of the FF.
It is not obvious that RESP charges provide the best description, especially due to its non-uniqe solution. 
Shall we solve RESP fitting with the constraint of full charges and then scale down, or shall we rather solve the fitting with a scaled total charge target?

acheivable accuracy of the MD engines themselves (mainly Hector's worry) is another limiting factor in fine tuning parameters -- solid physical ground helps.

application of the correction to other lipid models: The rule looks general, however, it depends on how accurate the original model was.
From preliminary simulations with POPE, it looks that the rule works for Lipid14 FF, at least for zwitterionic headgroups.

@@

Reiterate what we did...

@@

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Fig/OrderParameterIONSchangesSCALED.eps}
  \caption{\label{OrderParameterIONSchangesSCALED}
    Headgroup order parameter changes as a function of cations from models
    with modified headgroup partial charges.}
\end{figure}

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Fig/OrderParameterCHANGESnewMODELS.eps}
  \caption{\label{OrderParameterCHANGESnewMODELS}
    Headgroup order parameter changes as a function of cations from models
    with modified headgroup partial charges.}
\end{figure}

@@

\begin{table}
  \caption{Area per lipid from different models for POPC without ions\label{apls} }
  \begin{tabular}{c c}
    model          & A (Å$^2$)    \\
    Lipid14 (literature)  & 65.6$\pm$ 0.5   \\
    Lipid14ecc0.80+sigma0.875 & 64.3      \\
    Slipids (literature T=303K)       & 64.6 $\pm$ 0.4   \\
    SlipidsEEC0.8  & 57.8  \\
  \end{tabular}
\end{table}
